---
title: "MERS-RMarkdown"
author: "Madison"
date: "May 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#First we need to set the libraries and pull in the csv files

```{r mers}
library("lubridate")
library("ggplot2")
library(plotly)
getwd()
mers <- read.csv("cases.csv")
```

#The following will fix and then remove a line from our data
```{r pressure, echo=FALSE}
head(mers)
class(mers$onset)
head(mers$onset, 2)
mers$hospitalized[890] <- c('2015-02-20')
mers <- mers[-471,]
```

#Then we manipulize our data before we plot it
```{r}
mers$onset2 <- ymd(mers$onset)
mers$hospitalized2 <- ymd(mers$hospitalized)
class(mers$onset2)
day0 <- min(na.omit(mers$onset2))
mers$epi.day <- as.numeric(mers$onset2 - day0)
```

#This plot shows globals counts of MERS since the date of symptom onset
```{r plot}
ggplot(data=mers) + 
  geom_bar(mapping=aes(x=epi.day)) + 
  labs(x='Epidemic day', y='Case count' , title='Global count of MERS cases by date of symptom onset',
       caption="Data fom: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

#The next plot shows by country
```{r}
ggplot(data=mers) + 
  geom_bar(mapping=aes(x=epi.day, fill=country)) + 
  labs(x='Epidemic day', y='Case count' , title='Global count of MERS cases by date of symptom onset',
       caption="Data fom: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv") 
```

#Now we are altering by infectioud period onset date
```{r}
mers$infectious.period <- mers$hospitalized2-mers$onset2
class(mers$infectious.period)
mers$infectious.period <- as.numeric(mers$infectious.period, units= 'days')
```


#This plot is showing the new onset date
```{r}
ggplot(data=mers) +
  geom_histogram(aes(x=infectious.period)) +
  labs(x='Infectious period', y='Frequency', title='Distribution of calculated MERS infectious period',
       caption='Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv')
```



```{r}
mers$infectious.period2 <- ifelse(mers$infectious.period<0,0,mers$infectious.period)
```

#Histogram of changed infectious period date
```{r}
ggplot(data=mers) +
  geom_histogram(aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency', 
       title='Distribution of calculated MERS infectious perdiod (positive values only)', caption="Data from: //github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```


#Plot of the area of the frequency of infectious period date
```{r}
ggplot(data=mers) +
  geom_area(stat='bin', mapping=aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency', 
       title='Area plot for MERS infectious perdiod (positive values only)', caption="Data from: //github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```




#The plot below shows the difference in infectious period by epidemic day, while also showing the difference across gender and country
```{r}
ggplot(data=subset(mers, gender %in% c('M', 'F') & country %in% c('KSA', 'Oman', 'Iran', 'Jordan', 'Qatar', 'South Korea', 'UAE'))) +
  geom_point(mapping = aes(x=epi.day, y=infectious.period2, colour=country)) +
  facet_grid(gender ~ country) +
  scale_y_continuous(limits = c(0, 50)) +
  labs(x='Epidemic Day', y='Infectious Period', 
       title='MERS infectious period by gender and country', caption='Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv')
```



#The interactive plot below enables you to choose an epidemic day to view the country and number of cases at an exact period in time
```{r}
epi.curve <- ggplot(data=mers) +
  geom_bar(mapping=aes(x=epi.day, fill=country)) +
  labs(x='Epidemic day', y='Case count', title='Global count of MERS cases by date of symp' ,
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
ggplotly(epi.curve)
```







